---
title: "GWIP"
author: "Nate Breznau"
date: "`r Sys.Date()`"
output: html_document
---

Global Work-Injury Policy Database Longitudinal (GWIP_Long) v1

Sources:

Each individual indicator:

Welfare State Information Systems ([WeSIS](https://www.wesis.org/)) at the University of Bremen.

Complete dataset:

Harvard Dataverse

## Setup

```{r }

packages <- c("tidyverse",
              "pacman",
              "here",
              "readxl",
              "countrycode",
              "zoo")

pacman::p_load(packages, character.only = T)

```


## Data

### GWIP

```{r}

df_gwip <- read_csv(here("data", "gwip_long_v1.csv"))

df_cov <- df_gwip %>%
  mutate(iso3c = countrycode(country, "country.name", "iso3c"),
         cov = labor_workinjury_coverage_pct_lf,
         gen = as.numeric(labor_workinjury_replacement_rate_temp)) %>%
  select(iso3c, country, year, cov, gen)

# get colonial codes from gwip v3
df_col <- read_csv(here("data", "gwip_v3.csv")) %>%
  mutate(iso3c = countrycode(cow_code, "cown", "iso3c"),
         iso3c = ifelse(cow_code == 345, "SRB",
                 ifelse(cow_code == 817, "VNM",
                 ifelse(cow_code == 6, "PRI", iso3c)))) %>%
  select(iso3c, colonial_history) 

# global south codes
df_south <- read_csv(here("data", "global_south.csv")) %>%
  mutate(iso3c = countrycode(Country, "country.name", "iso3c"),
         global_south = South) %>%
  select(iso3c, global_south)

```

### Maddison

Bolt and Van Zanden (2024), "Maddison style estimates of the evolution of the world economy: A new 2023 update", Journal of Economic Surveys, 1–41

#### GDPpc

```{r}

df_mpd_gdp <- read_xlsx(here("data", "mpd2023_web.xlsx"), sheet = "GDPpc", skip = 2)

df_mpd_gdp <- df_mpd_gdp %>% filter(!is.na(year))

# Reshape long
df_gdp <- df_mpd_gdp %>%
  pivot_longer(
    cols = -year,
    names_to = "iso3c",
    values_to = "gdp"
  ) %>%
  filter(!is.na(gdp)) %>%
  as.data.frame()

```

#### Population

```{r}

df_mpd_pop <- read_xlsx(here("data", "mpd2023_web.xlsx"), sheet = "Population", skip = 2)

df_mpd_pop <- df_mpd_pop %>% filter(!is.na(year))

df_pop <- df_mpd_pop %>%
  pivot_longer(
    cols = -year,
    names_to = "iso3c",
    values_to = "pop"
  ) %>%
  filter(!is.na(pop)) %>%
  as.data.frame()

```

## Full Data

### Combine

Ensure that there is a row for every 10 years for every country since 1880

```{r}

df <- df_cov %>%
  full_join(df_gdp, by = c("iso3c", "year")) %>%
  full_join(df_pop, by = c("iso3c", "year")) 

years <- seq(1880, 2020, by = 10)

df_decades <- expand.grid(
  iso3c = unique(df$iso3c),
  year = years
)

df <- full_join(df_decades, df, by = c("iso3c", "year"))

df <- df %>%
  left_join(df_col, by = c("iso3c")) %>% 
  left_join(df_south, by = c("iso3c")) %>%
  arrange(iso3c, year) 

```

### Interpolate

Linear interpolation between non-missing values for pop and gdp

Last observation carried forward for cov and gen

```{r}

# remove countries with no GWIP data

df <- df %>%
  # Change -99 to NA
  mutate(across(where(is.numeric), ~ na_if(., -99))) %>%
  
  #Drop countries with missing cov and gen
  group_by(iso3c) %>%
  filter(!(all(is.na(cov)) & all(is.na(gen)))) %>%
  ungroup() %>%
  as.data.frame()

df <- df %>%
  group_by(iso3c) %>%
  arrange(year) %>%
  mutate(
    pop_i = na.approx(pop, x = year, na.rm = FALSE),
    gdp_i = na.approx(gdp, x = year, na.rm = F)
  ) %>%
  arrange(iso3c, year) %>%
  ungroup() %>%
  as.data.frame()

df <- df %>%
  group_by(iso3c) %>%
  arrange(year) %>%
  mutate(
    first_nonmissing_year = min(year[!is.na(cov)], na.rm = TRUE),
    cov_i = if_else(
      year < first_nonmissing_year, 
      NA_real_,  # Keep NA before first observation
      na.locf(cov, na.rm = FALSE)  # Carry forward after first observation
    )
  ) %>%
  select(-first_nonmissing_year) %>%  # Clean up
  ungroup()

df <- df %>%
  group_by(iso3c) %>%
  arrange(year) %>%
  mutate(
    first_nonmissing_year = min(year[!is.na(gen)], na.rm = TRUE),
    gen_i = if_else(
      year < first_nonmissing_year, 
      NA_real_, 
      na.locf(gen, na.rm = FALSE)  
    )
  ) %>%
  select(-first_nonmissing_year) %>%  # Clean up
  ungroup() %>%
  arrange(iso3c, year) %>%
  as.data.frame()

# fill in missing country names
df <- df %>%
  mutate(country = countryname(iso3c, "iso3c", "country.name"))

```

### Save

```{r}
write.csv(df, here("data", "df.csv"), row.names = F)
```


## Citations

```{r cites}

packages %>%
  map(citation) %>%
  print(style = "text")

```

## Colophon

```{r colo}

sessionInfo()

```

